---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

TileManager: Tools for creating and detecting tiling schemes for geospatial datasets
---------------

**Authors:** Andrew Plowright<br/>
**License:** GPL 3

[![Build Status](https://travis-ci.org/AndyPL22/TileManager.svg?branch=master)](https://travis-ci.org/AndyPL22/TileManager)

```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "READMEimgs/",
  fig.height = 4,
  fig.width = 6
)

knitr::opts_knit$set(
  global.par = TRUE
  )
```

```{r, echo = FALSE}
par(mar = c(3,3,1,0.1))
```

This package provides tools for working with tiled geospatial datasets.

#### Introduction

Use the `tileScheme` function to create a set of tiles from a Raster or Extent object.
```{r demo1}
library(TileManager)

data(CHMdemo)

ts <- tileScheme(CHMdemo, tiledim = c(30,30), buffer = 5)

plot(CHMdemo)
plot(ts, add = T)

```

Use the `removeEmpty` argument to drop tiles with no Raster values.
```{r demo2}
ts <- tileScheme(CHMdemo, tiledim = c(20,20), buffer = 2.5, removeEmpty = TRUE)

plot(CHMdemo)
plot(ts, add = T)

```

Other handy features:

* The `origin` argument can be used to force the tile scheme to snap to a pair of coordinates.
* The `bufferSpill` argument controls whether or not the buffers extent beyond the input's original extent.
* By default, tile dimensions are in map units. Using the `cells` argument, they can be defined by a number of Raster cells.


#### The 'tileScheme' class

A 'tileScheme' object is composed of the following slots:
```{r}
str(ts, 2)
```

* Three lists of Polygons objects:
    + `tiles`: the actual extents of the tiles
    + `buffs`: the buffered tiles
    + `nbuffs`: the non-overlapping buffers (see section below)
* The `crs` slot, which stores the tile scheme's coordinate reference system.
* A numeric `buffer` slot.
* A data.frame in the `data` slot, which stores the row, column and name of each tile.
```{r}
head(ts@data)
```

#### Non-overlapping buffers

Non-overlapping buffers (often abbreviated to **nbuffs**) are useful for re-assembling tiled data. Essentially, they preserve buffers only where they _do not overlap onto neighboring tiles_ (i.e.: along the edge of the tile scheme). This allows you to recombine tiles without worrying about overlapping areas _and_ without losing any information along the data edges.

In the example below:

* The __solid blue__ is the tile
* The __dashed red__ is the buffer
* The __solid red__ is the non-overlapping buffer
```{r nbuff-sample, echo = FALSE}
plot(CHMdemo, xlim = c(439871, 439960), ylim = c(5526705, 5526756), xlab = NA, ylab = NA, legend = FALSE)

plot(ts, add = T)
```

#### Methods

Some useful methods are provided for subsetting the tile scheme, or for converting it into other formats.

Get buffers as a SpatialPolygonsDataFrame:

```{r}
ts[["buffs"]]
```

Subset a specific tile by name, number or row/col:

```{r}
# By name
ts["R2C2"]

# By number
ts[7]

# By row/col
ts[2,3]
```

Subset entire rows or columns:

```{r}
# One row
ts[4,]

# Multiple columns
ts[,2:3]
```

#### Saving and loading

The tile scheme can be saved as a single SHP file. In this case, `tiles`, `buffs` and `nbuffs` will all be merged into a single polygons dataset. Buffer information is saved to an accompanying XML file.

```{r, eval = FALSE}
# Create tile scheme
ts <- tileScheme(CHMdemo, tiledim = c(30,30), buffer = 5)

# Save tile scheme
tileSave(ts, "C:/myfiles/tilescheme.shp")

# Load tile scheme
ts <- tileLoad("C:/myfiles/tilescheme.shp")
```


#### Tile detection

If you have a received a series of tiled raster files, the `tileDetector` function can even be used to detect the tile size and buffer size of the data and generate the associated `tileScheme`.
